Portfolio prioritization of salmon conservation

### 2013-03-28 inst/examples/salmonport.tex
Update all remaining files

% my modifications:
\usepackage{fontspec}
\usepackage{booktabs}
\usepackage{mathptmx}
\defaultfontfeatures{Ligatures=TeX}
\setsansfont{Calibri}
\setmonofont{Inconsolata}
%\setmainfont{Linux Libertine}
%\setmainfont{Times New Roman}
\setmainfont{Calibri}
%\setmainfont{Helvetica}
\DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }
\widowpenalty=1000
\clubpenalty=1000
\usepackage{setspace}
\onehalfspacing
%\doublespacing

@@

I've implemented straying as in \citet{Cooper1999}. I generate a matrix
that represents the fraction of straying between any two populations. I
have set up the metapopulation (thus far) in a very simple scenario: the
populations are arranged in a line and those that are nearer to each
other are more likely to stray between each other {[}insert caveats{]}.
Two parameters control the straying: the fraction of fish $f$ that stray
from their natal stream in any given generation and the rate at which
this straying between streams decays with distance $m$. I calculated the
number of salmon straying from stream $j$ to stream $i$ as,

\begin{equation}
  \mathrm{strays}_{tij} = f R_{tj} \frac{e^{-m \lvert i-j \rvert }}{\displaystyle\sum\limits_{\substack{k = 1 \\
  k \neq j}}^{n} e^{-m \lvert k-j \rvert }},
\end{equation}

@@

\begin{figure}[htbp]
\centering
\includegraphics{figure/thermal-curves.pdf}
\caption{Thermal tolerance curves for ten populations. Here, and
throughout this document, I have used the same rainbow of colours to
identify the salmon populations. The populations are arranged from blue
(more productive with cooler temperatures) to red (more productive with
warmer temperatures).\label{fig:thermal-curves}}
\end{figure}

@@

The following are options and their default values:

\begin{verbatim}
n_t = 120, # number of generations
n_pop = 10, # number of populations
stray_decay_rate = 0.3, # rate that straying decays with distance
stray_fraction = 0.01, # fraction of fish that stray from natal streams
b = rep(1000, n_pop), # Ricker density-dependent parameter
spawners_0 = round(b), # spawners at start
sigma_v = 0.1, # stock-recruit residual SD
v_rho = 0.4, # stock-recruit residual AR1 correlation
a_width_param = rep(0.02, n_pop), # width of thermal curves by pop
optim_temp = seq(13, 19, length.out = n_pop), # optimal temperatures by pop
max_a = rep(1.4, n_pop), # maximum Ricker a values by pop at optimum temp
env_type = c("sine", "arma", "regime", "linear", "constant"),
env_params = list(amplitude = 2.0, ang_frequency = 0.2, phase = 0, mean_value = 16),
use_cache = FALSE, # regenerate stochastic values? 
add_straying = TRUE # include or ignore straying
\end{verbatim}

First, let's run a couple example simulations:

\begin{figure}[htbp]
\centering
\includegraphics{figure/run-base-example1.pdf}
\caption{Example simulation with all parameters at default values. I'm
only showing a sample of 40 generations (after the burn-in period). From
top to bottom: salmon returns each generation (catch plus escapement);
Ricker $a$ parameter; the environmental signal (could think of it as
temperature or a composite environmental signal like the PDO); number of
salmon straying by population; number of straying salmon being received
by population; fisheries catch based on escapement MSY calculation;
salmon escapement; residuals from the spawner-return curve.}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{figure/plot-various-options-ts.pdf}
\caption{Example simulations with different parameter values. This
version has ten populations.\label{fig:sim-param-ts-10}}
\end{figure}

@@

@@

>>>>

% my modifications:
\usepackage{fontspec}
\usepackage{booktabs}
\usepackage{mathptmx}
\defaultfontfeatures{Ligatures=TeX}
\setsansfont{Calibri}
\setmonofont{Inconsolata}
\setmainfont{Linux Libertine}
%\setmainfont{Times New Roman}
%\setmainfont{Calibri}
%\setmainfont{Helvetica}
\DeclareTextCommandDefault{\nobreakspace}{\leavevmode\nobreak\ }
\widowpenalty=1000
\clubpenalty=1000
\usepackage{setspace}
\onehalfspacing
%\doublespacing

@@

I've implemented straying as in \citet{Cooper1999}. I generate a matrix
that represents the fraction of straying between any two populations. I
have set up the metapopulation (thus far) in a very simple scenario: the
populations are arranged in a line and those that are nearer to each
other are more likely to stray between each other {[}insert caveats{]}.
Two parameters control the straying: the fraction of fish $f_{stray}$
that stray from their natal stream in any given generation and the rate
at which this straying between streams decays with distance $m$. I
calculated the number of salmon straying from stream $j$ to stream $i$
as,

\begin{equation}
  \mathrm{strays}_{tij} = f_{stray} R_{tj} \frac{e^{-m \lvert i-j \rvert }}{\displaystyle\sum\limits_{\substack{k = 1 \\
  k \neq j}}^{n} e^{-m \lvert k-j \rvert }},
\end{equation}

@@

\begin{figure}[htbp]
\centering
\includegraphics{figure/thermal-curves.pdf}
\caption{Thermal tolerance curves for ten populations. Here, and
throughout this document, I have used the same rainbow of colours to
identify the salmon populations. The populations are arranged from blue
(more productive with cooler temperatures) to red (more productive with
warmer temperatures). The vertical lines indicate mean, minimum, and
maximum values for one scenario --- you can ignore them for
now.\label{fig:thermal-curves}}
\end{figure}

\section{Summary of simulation parameters}

\begin{table}[h!]
\centering
\small
\caption{Salmon metapopulation parameters with base case and alternate values.}
\begin{tabular}{p{7.0cm}p{1.6cm}p{3.2cm}}
\toprule
Description & Parameter & Base [lower, upper] \\
\midrule
Stock-recruit residual standard deviation (on log scale) & $\sigma_v$ & 0.30 [0.05, 0.50] \\
First order (AR1) serial correlation of stock-recruit residuals & $\rho_w$ & 0.40 [0, 0.80] \\
Fraction of fish that stray from natal streams & $f_{stray}$ & 0.02 [0, 0.10] \\
Exponential rate of decay of straying with distance & $m$ & 0.30 [0.05, 0.50] \\
Standard deviation of beta distribution for implementation error & $\sigma_{impl}$ & 0.05 [0, 0.20] \\
Frequency of assessment (years) & $f_{assess}$ & 20 [5, 50] \\
\bottomrule
\end{tabular}
\label{tab:salm-pars}
\end{table}

@@

\begin{figure}[htbp]
\centering
\includegraphics{figure/run-base-example1.pdf}
\caption{Example simulation with all parameters at default values. I'm
only showing a sample of 40 generations (after the burn-in period). From
top to bottom: salmon returns each generation (catch plus escapement);
Ricker $a$ parameter; the environmental signal (could think of it as
temperature or a composite environmental signal like the PDO); number of
salmon straying by population; number of straying salmon being received
by population; fisheries catch based on escapement MSY calculation;
salmon escapement; residuals from the spawner-return curve; estimated
Ricker $a$ parameter; estimated Ricker $b$ parameter.}
\end{figure}

@@

@@

\begin{verbatim}
Error: ignoring SIGPIPE signal
\end{verbatim}

@@

\begin{figure}[htbp]
\centering
\includegraphics{figure/plot-cons-plans-curves.pdf}
\caption{Different ways of prioritizing conservation. The curves show
the environmental-tolerance curves for the populations that are
conserved. The colours are the same as earlier and identify the
different populations. These are in the same order as they are referred
to in the next two plots: full range of environmental response
diversity, most stable only, one half, other half, most asynchronous.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plot-cons-plans-mv.pdf}
\caption{Conservation plans in mean-variance space. The dots show
portfolios from individual simulation runs. The dark and light shaded
areas show 25\% and 75\% quartiles. The left panel shows a scenario with
a stationary environmental signal and strong short-term environmental
noise. The right panel shows a scenario with a linearly-increasing
environmental signal and minimal short-term environmental noise. These
two scenarios bound the two environmental extremes: strong short-term
noise with minimal long-term signal vs.~a strong long-term signal with
minimal short-term signal. Notice what happens to the two \emph{half}
conservation strategies. A more realistic scenario combines the two. The
efficient frontier is shown for the displayed portfolios. This isn't
necessarily the overall efficient frontier.}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics{figure/plot-cons-plans-ts.pdf}
\caption{Conservation plans in terms of metapopulation abundance. One
realization with a periodic (sine wave) environmental signal is shown as
an example. The light-grey lines in the background show 100
randomly-sampled portfolios with the same stock-recruit residuals and
the same environmental signal. Notice how much less variable the
scenarios with balanced response diversity are. The two scenarios with
unbalanced response diversity tend to seesaw with the environment.}
\end{figure}

\section{Take home messages}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  If the dominant environmental signal is a long-term shift, and you
  keep around one half of the response diversity, then you get different
  portfolio means but the same variances.
\item
  If the dominant environmental signal is short-term, and you keep
  around one half of the response diversity, then you get different
  portfolio variances but the same means.
\item
  Combine those two (which best reflects reality) and you get both
  different means and variances.
\item
  Whether you invest in just stable, just asynchronous, or a mix of
  both, all that really matters is that you have as many populations as
  possible and as balanced as possible response diversity.
\item
  An even weighting across as many possible streams (assuming balanced
  response diversity) gets you pretty close to the minimum-variance
  portfolio. {[}not shown currently{]}
\item
  With systematic long-term environmental change that dominates: if you
  can forecast the environment and know the responses of salmon
  populations from individual streams then you can increase your
  expected rate of metapopulation growth with minimal effect on variance
  by prioritizing the populations that will do well. Although if you get
  it wrong you will do worse than if you had hedged your bets and
  conserved response diversity.
\item
  But, if there are also strong short-term environmental fluctuations
  (and there probably are), then there's always a benefit to portfolio
  variance by having some response diversity.
\item
  Variable stocks aren't inherently undesirable to conserve, as long as
  they are balanced or averaged out.
\item
  Take home message: ``To keep every cog and wheel is the first
  precaution of intelligent tinkering.'' But, if you must prioritize,
  keep a selection of different cogs and wheels --- especially a
  spatially diverse set of cogs and wheels. That gives you the best
  chance of an efficient ecological portfolio.
\end{enumerate}

\section{Open questions and next steps}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  How do smaller or larger values of the various parameters affect the
  results?
\item
  Are the results sensitive to the height and width of the thermal
  tolerance curves? (I think within reason, no.)
\item
  Are these results robust to, say, cyclical or regime shift style
  environmental noise?
\item
  Should this be a generic ``environmental signal'' or temperature?
\item
  Show what happens for a given scenario as you conserve more streams
  (it gets closer to the minimum-variance portfolio).
\item
  Show an example of where these conservation plans place you in
  mean-variance space across all possible weights. (Done but not shown.)
\item
  Bring these conservation plans back to established suggestions
  (e.g.~VSP).
\item
  How can we simplify the main figures?
\item
  Is there a way to combine the main figures to make the result more
  understandable (e.g.~little response-diversity curves as a legend)?
\item
  Redo the Monte Carlo portfolios to show an extreme example where you
  can see the weights shifting. For example, a linearly increasing
  environmental signal with some generation-to-generation noise.
\item
  Generations or years? Years seem simpler to talk about. Generations
  seem more widely applicable and accurate.
\end{enumerate}
